For Induction-Induction, Induction is Enough

### 2019-11-10 paper/paper.tex
section 4

\begin{document}

\maketitle

%TODO mandatory: add short abstract of the document
\begin{abstract}
  Inductive-inductive types (IITs) are a generalisation of inductive types in
  type theory. They allow the mutual definition of types with multiple sorts
  where later sorts can be indexed by previous ones. An example is the James
  Chapman-style syntax of type theory with conversion relations for each sort
  where e.g.\ the sort of types is indexed by contexts.  In previous work it was
  shown that all finitary IITs can be constructed from a quotient
  inductive-inductive type (QIIT), namely the theory of IIT signatures. This is
  a small domain-specific type theory where a context is a signature for an
  IIT. In this paper we construct this theory of signatures only using inductive
  types. We rely on function extensionality and uniqueness of identity
  proofs. All of the constructions of this paper were formalised in the proof
  assistant Agda.
\end{abstract}

@@

\section{Constructing IITs from the Theory IIT of Signatures}
\label{sec:andras}

Explain that the previous paper was modular.

>>>>

\begin{document}
\maketitle

\begin{abstract}
  Inductive-inductive types (IITs) are a generalisation of inductive types in
  type theory. They allow the mutual definition of types with multiple sorts
  where later sorts can be indexed by previous ones. An example is the James
  Chapman-style syntax of type theory with conversion relations for each sort
  where e.g.\ the sort of types is indexed by contexts.  In previous work it was
  shown that all finitary IITs can be constructed from a quotient
  inductive-inductive type (QIIT), namely the theory of IIT signatures. This is
  a small domain-specific type theory where a context is a signature for an
  IIT. In this paper we construct this theory of signatures using only inductive
  types, thereby showing a reduction of all finitary IITs to inductive types.
  We rely on function extensionality and uniqueness of identity proofs. All
  constructions in this paper were formalised in the proof assistant Agda.
\end{abstract}

@@

\section{Constructing IITs from the Theory of IIT Signatures}
\label{sec:andras}

We construct all IITs describable by the theory of IIT signatures through the
term model construction of \cite{Kaposi:2019:CQI:3302515.3290315}. There, it is
shown that from the syntax of the theory of signatures for finitary QIITs, one
can construct any particular finitary QIIT. The idea is that for a signature
$\Gamma$, the initial algebra can be built from sets of terms of the form
$\Tm\,\Gamma\,(\El\,a)$.

In this section, we only need to check that the signatures in
\cite{Kaposi:2019:CQI:3302515.3290315} and the term model construction restricts
to IITs, i.e.\ that from a syntax for the theory of IIT signatures all finitary
IITs are constructible.

\begin{itemize}
\item
  Finitary IIT signatures are obtained exactly by dropping equality constructors from
  finitary QIIT signatures. It follows that any model for the theory of QIIT signatures
  restricts to IITs, hence we inherit the categorical semantics of QIITs.
\item
  The term model construction also restrict in a straightforward way. For QIITs,
  term algebras are constructed by induction on the syntax of signatures, and
  then another induction constructs the eliminator (i.e.\ dependent induction
  principle). In both cases, simply dropping equality constructors from the
  induction yields restriction to IITs.
\end{itemize}

However, we shall mention that the Agda formalisation for the current paper
cannot be directly plugged into the Agda code for
\cite{Kaposi:2019:CQI:3302515.3290315} (which includes most of the term model
construction). One reason is that the QIIT formalisation uses strict computation
rules for induction over signatures, while here we show propositional
computation. This mismatch can be in principle remedied by noting that both
formalisations use UIP and function extensionality, hence by the known
translations between extensional and intensional type
theories\cite{hofmann95extensional,winterhalter2019eliminating}, we can switch
between strict and propositional equalities.

Also, the current paper proves initiality, i.e.\ unique recursion for
signatures, while \cite{Kaposi:2019:CQI:3302515.3290315} uses dependent
elimination. We expect that the two notions are equivalent. An extension of
\cite{Kaposi:2019:CQI:3302515.3290315} to large infinitary QIITs would entail
this equivalence, since that would cover the (large, infinitary) theory of IIT
signatures.
